---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rxs2xcell

<!-- badges: start -->
<!-- badges: end -->

The goal of `rxs2xcell` is to prepare, quality check and harmonize a collection of
QWA images and corresponding ROXAS output files for submission to the QWA database.

## Installation

You can install the development version of `rxs2xcell` from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("mnaegelin/rxs2xcell")

# or:
remotes::install_github('mnaegelin/rxs2xcell', auth_token = GITHUB_PAT)

# TODO: change to final version under organization in future
```

## Usage

Use the package to extract all available metadata from the ROXAS output files under a given path.

```{r example, eval = FALSE}
library(rxs2xcell)

# set input and output paths and dataset name
path_in <- '../QWA_Arzac2024'
path_out <- './out'
dataset_name <- 'QWA_Arzac2024'


# get overview of data to be read and extract data structure from filenames
files <- get_input_files(path_in)
df_structure <- extract_data_structure(files)


# read available metadata
df_meta <- collect_metadata_from_files(df_structure,
                                       roxas_version='classic')

# write extracted metadata to file
write.csv(df_meta, file.path(path_out, paste0(dataset_name, '_meta_extracted.csv')),
          row.names = FALSE)

# TODO:
# maybe write a use_template() function that takes a template file and opens it, so no copy paste needed?
# maybe just a wrapper around:
# usethis::use_template('template_workflow.Rmd', package='rxs2xcell')
# maybe something like the use_targets() function from https://docs.ropensci.org/targets/reference/use_targets.html
```


To provide the additional metadata required (e.g. site and tree information, author details, ...) for the submission of a dataset, 
you can use the interactive Shiny app which is part of this package.

The app is hosted on a WSL server and accessible via [LINK](https://shiny16.wsl.ch:3838/naegelin/xcell_meta). 

Or, if you want to run it locally, you can use the following command:

```{r, eval = FALSE}
launch_metadata_app()
```



To load and preprocess the cell and ring measurements:

```{r, eval = FALSE}
# read raw cells/rings data
QWA_data <- collect_raw_data(df_structure)

# clean raw data
QWA_data <- validate_QWA_data(QWA_data, df_meta)

# remove outliers
# NOTE: ROXAS does some threshold based outlier checks and assigns these a
# a negative value. Here, we replace these outliers with NAs
QWA_data <- remove_outliers(QWA_data)

# complete cell measures
QWA_data <- complete_cell_measures(QWA_data)

# save the cleaned data to files
write.csv(QWA_data$cells,
          file.path(path_out, paste0(dataset_name, '_cells.csv')),
          row.names = FALSE)

write.csv(QWA_data$rings,
          file.path(path_out, paste0(dataset_name, '_rings.csv')),
          row.names = FALSE)

```



To check the coverage and add ring flags:
```{r, eval = FALSE}
launch_coverage_app()

# TODO: make it so ring flags overwrite files?
```



